---
title: "Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Starting

The package `tktdjl2r` is a companion package to use from `R` the package `TKTDsimulations.jl` written in `julia`.

`TKTDsimulations.jl` is dedicated to the simulation of a large set of TKTD models using efficient differential equations solvers
implemented in `julia`.

You need to have `julia` installed in your computer:[see here for information about julia installation](https://julialang.org/downloads/)

If `Julia` is on your machine, then install and load `tktdjl2r` library:

```{r setup, eval=FALSE}
library(tktdjl2r)
```

In order to make the link from `julia` to `R`, you have to run the setup command:

```{r setup_tktdjl2r, eval=FALSE}
# Julia can be quite long to install the first time.
tktdjl2r::tktdjl2r_setup()
```

This command is going to call the `JuliaCall` R package and check or install required dependencies.

In the console, you should have (for instance on windows):

```
Julia version 1.5.0 at location C:\Users\...\AppData\Local\Programs\JULIA1~1.0\bin will be used.
Loading setup script for JuliaCall...
Finish loading setup script for JuliaCall.
```

# Toxicokinetics

The first time this code is run can be long. But after the first run, the run should be highly faster.

```{r runTK, eval=FALSE}
single_runTK = runTK(c(0,1,2,3), c(0,1,2,2), 0.5)
```


```{r runTK_MCMC, eval=FALSE}
kdVector = rbeta(4,10,10)

mcmc_runTK = runTK_MCMC(c(0,1,2,3), c(0,1,2,2), kdVector)
```


## TKTD, Stochastic Death type SD

```{r runSD, eval=FALSE}
single_runSD = runSD(c(0,1,2,3), c(0,1,2,2), 0.5, 0.2, 1, 0.4)
```


```{r runSD_MCMC, eval=FALSE}
paramDF = data.frame(
  kd = rbeta(4,10,10),
  hb = rbeta(4,10,10),
  z = rbeta(4,10,10),
  kk = rbeta(4,10,10))
mcmc_runSD = runSD_MCMC(c(0,1,2,3), c(0,1,2,2), paramDF$kd, paramDF$hb, paramDF$z, paramDF$kk)
```

## TKTD, Individual Tolerance type IT

```{r runIT, eval=FALSE}
single_runIT = runIT(c(0,1,2,3), c(0,1,2,2),0.5, 0.2, 1, 0.4)
```

```{r runIT_MCMC, eval=FALSE}
paramDF = data.frame(
  kd = rbeta(4,10,10),
  hb = rbeta(4,10,10),
  alpha = rbeta(4,10,10),
  beta = rbeta(4,10,10))
mcmc_runIT = runIT_MCMC(c(0,1,2,3), c(0,1,2,2), paramDF$kd, paramDF$hb, paramDF$alpha, paramDF$beta)
```

# Comparison with pacakge 'deSolve'

## Test for toxicokinetics, TK

```{r deSolveTK, eval=FALSE}
library(deSolve)

model_TK <- function(t, State, parms, input) {
  with(as.list(c(parms, State)), {
    list(kd*(input(t) - State))    # internal damage
  })
}

deSolve_TK <- function(time, conc, listParameters){
  signal <- data.frame(times = time,
                       import = conc)

  sigimp <- stats::approxfun(signal$times,
                             signal$import,
                             method = "linear",
                             rule = 2)

  ## values for steady state
  xstart <- c(D = 0)
  ## model
  out <- ode(y = xstart,
             times = time,
             func = model_TK,
             parms = listParameters,
             input = sigimp)

  data.frame(
    time = time,
    exposure = conc,
    TK = out[,2]
  )

}
```

### Plot both

```{r plotTK, eval=FALSE}
listParameters = list(kd = 0.5)
testR_runTK = deSolve_TK(c(0,1,2,3), c(0,1,2,2), list(kd=0.5))
plot(c(0,1,2,3), single_runTK$TK, type = "l", lwd = 3)
lines(c(0,1,2,3), testR_runTK$TK, col = "red", lwd = 2)
```

### Test speed

```{r speedTK, eval=FALSE}
library(microbenchmark)

microbenchmark::microbenchmark(
  testJulia_runTK = runTK(c(0,1,2,3), c(0,1,2,2), 0.5),
  testR_runTK = deSolve_TK(c(0,1,2,3), c(0,1,2,2), list(kd=0.5)),
  times = 10
)
```


